### 2021-07-28 MSA_Notum_journey.Rmd
add iqtree and iqtree results with MUSCLE msa

---
title: "The journey of Multiple Sequence Alignment"
subtitle: "Diving in the Notum"
author: "Constantinos Yeles (Konstantinos Geles)"
date: "First version: Tue 26/07/2021, Latest update: `r format(Sys.time(), '%a %d/%b/%Y')`"
output:
  html_document:
    toc: yes
    toc_depth: 3
    theme: paper 
  pdf_document:
    toc: yes
    toc_depth: 3
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

# Blast alignment
First we need to download the fasta sequence and the database from
dropbox? Maybe can be uploaded later to zenodo 
(probably I will have more context to add later).

We will use local blast run on docker as the [link](https://faculty.virginia.edu/wrpearson/fasta/) wasn't working.

```{bash}
# get the docker and run
docker pull ncbi/blast
docker run --rm -ti -v $(pwd):/home/my_data  ncbi/blast
# move inside /home dir of the container
cd /home

# create a folder for the ouput of the blast database and unzip the fasta sequences
mkdir my_data/some_species_blast_db
gunzip my_data/oma_db.gz

# make a blast database 
makeblastdb  -dbtype "prot" -in "my_data/oma_db" -input_type "fasta" -out "my_data/some_species_blast_db/blast_oma.db" -logfile "my_data/some_species_blast_db/logfile"

# blast the sequence of Notum
blastp -query "my_data/dm_notum.fasta"  -db "my_data/some_species_blast_db/blast_oma.db" -out "my_data/results_blast.txt" -outfmt "7 delim=@ qseqid qlen sseqid slen qstart qend sstart send qseq sseq evalue bitscore score length pident nident mismatch positive ppos frames qframe sframe sstrand qcovs qcovhsp qcovus" -num_threads 4
```

# Import results into R
```{r include=FALSE, eval = TRUE}
# load libraries
suppressPackageStartupMessages({
library(tidyverse)
library(plyranges)
library(vroom)
  })

# function to import the results
read_blast_tab_format <- function(blast_file, delim = "@"){
  fourth_line_fields <- blast_file %>% 
  vroom_lines() %>% 
  .[4] %>% 
  str_remove("# Fields: ") %>% 
  str_split(", ") %>% 
  unlist %>% 
  str_replace_all(" |/","_") %>% 
  str_replace("%","percent") %>% 
  str_remove("\\.")

  blast_sequences <- blast_file %>% 
  read_delim(delim = delim, # here it should be made clear that the delim could be something else
             comment = "#", 
             col_names = fourth_line_fields) %>% 
  mutate(strand = ifelse(subject_strand == "minus","-","+") %>% as_factor()) %>% 
  mutate(
         start = ifelse(strand == "-", yes = s_end, no = s_start),
         end = ifelse(strand == "-", yes = s_start,no = s_end),
         seqnames = subject_id) %>% 
    as_granges()

  return(blast_sequences)
}
# import the results

notum_blst <- read_blast_tab_format(blast_file = "results_blast.txt")
```

## print an example of the results 
```{r eval = TRUE}
notum_blst %>% arrange(evalue)
```

## get the 10 top hits identifiers
```{r eval = TRUE}
top10_seqs <- notum_blst %>% arrange(evalue) %>% head(11) %>% .$subject_id
```

## import fasta file with all sequences
import the sequences and filter them for the 10 best hits from the previous results in order to make the multiple sequence alignments
```{r eval = TRUE}
suppressPackageStartupMessages({
library(Biostrings)
  })
fasta_db <- readAAStringSet("oma_db")

# I didn't consider that the names will be cut in the db creation thus I have to manipulate them to work properly for the filtering
names_fst_db <- fasta_db  %>% names %>% str_remove(" .+")

# filter for the top 10
fasta_db_fil <- fasta_db[names_fst_db %in% top10_seqs]
```

# Multiple Sequence Alignment
I found two packages in R implementing various algorithms. 
One is [msa](https://www.bioconductor.org/packages/release/bioc/html/msa.html) and the other [DECIPHER](https://www.bioconductor.org/packages/release/bioc/html/DECIPHER.html)
for now I will try msa
```{r eval = TRUE}
suppressPackageStartupMessages({
library(msa)
library(bios2mds)
  })
myFirstAlignment <- msa(fasta_db_fil)
```

## print the MSA
```{r eval = TRUE}
print(myFirstAlignment, show="complete")
```

## get significant sequences
```{r eval = TRUE}
sign_seqs <- notum_blst %>% filter(evalue < 0.05) %>% .$subject_id

# filter for the significant sequences
fasta_db_fil <- fasta_db[names_fst_db %in% sign_seqs]
```

## perform MSA with MUSCLE
```{r}
sign_seqs_muscle <- msaMuscle(fasta_db_fil)
sign_seqs_bio2mds <- msa::msaConvert(sign_seqs_muscle, type = "bios2mds::align")
export.fasta(sign_seqs_bio2mds, outfile = "sign_seqs_msa_muscle.fa", ncol = 60, open = "w")
```
check the sign_seqs_msa_muscle.fa file for the alignments

# IQTREE
```{bash}
docker run --rm -ti -v $(pwd):/home/my_data staphb/iqtree

cd home
mkdir my_data/my_first_tree

iqtree -s my_data/sign_seqs_msa_muscle.fa -nt 6 -mset LG -pre my_data/my_first_tree/my_tree

```